### 2023-04-06 appendix/acknowledgements.tex
Initiate PhD thesis structure

I am extremely grateful to my advisor, Sasha Aravkin, for providing support, guidance, and enthusiasm during my work on this project in the Department of Applied Mathematics and in Institute for Health Metrics and Evaluation.\\
I would also like to thank Jim Burke for his detailed review of my work which would hopefully lead to fruitful collaboration. \\ 
Last, but not least: I am grateful to Damian Santomauro, who introduced me to his work on the consequences of bullying, provided with a dataset, and with a valuable feedback on the performance of the proposed method. \\