### 2023-04-19 phd_thesis.tex

\documentclass[12pt,a4paper]{book}

% PRESETS OF PACKAGES AND MATH %
\input{latex-macros/useful_packages}
\input{math_macros-jim}

\usepackage{natbib}
% Color 
\newcommand{\blue}[1]{\textcolor{blue}{#1}}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\plum}[1]{\textcolor{Plum}{#1}}
\newcommand{\teal}[1]{\textcolor{teal}{#1}}

%%%%% USER CONFIGURATION %%%%%
\newcommand{\userName}{Aleksei Sholokhov}
\newcommand{\userId}{aksh}
\newcommand{\department}{Department of Applied Mathematics}
\newcommand{\institution}{University of Washington}
\newcommand{\projectNameShort}{}
\newcommand{\doctitle}{Doctoral Thesis}
\newcommand{\projectNameLong}{Optimization Methods for Parameter Identifications \\ in Settings with Only Partial Knowledge}

% Cast
\newcommand{\Peng}[1][]{\textbf{Peng\ifthenelse{\equal{#1}{}}{ }{ (#1):}}}
\newcommand{\Sasha}[1][]{\textbf{Sasha\ifthenelse{\equal{#1}{}}{ }{ (#1):}}}
\newcommand{\Lesha}[1][]{\textbf{\userId\ifthenelse{\equal{#1}{}}{ }{ (#1):}}}
\newcommand{\Brad}[1][]{\textbf{Brad\ifthenelse{\equal{#1}{}}{ }{ (#1):}}}
\newcommand{\Jim}[1][]{\textbf{Jim\ifthenelse{\equal{#1}{}}{ }{ (#1):}}}
\newcommand{\ouralgo}{MSR3 }

% Commands which are specific for this document 
\DeclareMathOperator{\bBeta}{\bar{\beta}}
\DeclareMathOperator{\bgamma}{\bar{\gamma}}
\DeclareMathOperator{\bSigma}{\bar{\Sigma}}

%%%%%%%%%%%%%%%
% PAGE FORMAT %
%%%%%%%%%%%%%%%

\pagestyle{fancy}
\setlength\parindent{0in}
\setlength\parskip{0.1in}
\setlength\headheight{15pt}
\linespread{1.18}

%%%%%%%%%%% HEADER / FOOTER %%%%%%%%%%%
\chead{\textit{\projectNameShort}}
\lhead{\textsc{\userName}}
\rhead{\textsc{\doctitle}}
\rfoot{}
\cfoot{\color{gray} \textsc{\thepage~/~\pageref*{LastPage}}}
\lfoot{}

% University Logo
%\newcommand{\univlogo}{%
%  \noindent % University Logo
%  \begin{wrapfigure}{r}{0.25\textwidth}
%    \vspace{-24pt}
%    \begin{center}
%      \includegraphics[width=0.25\textwidth]{Images/univ-logo.jpg}
%    \end{center}
%    \vspace{-10pt}
%  \end{wrapfigure}
%}

%\renewcommand{\thesection}{\Roman{section}}
% \renewcommand{\thesubsection}{\thesection.\Roman{subsection}}

%%%%%%%%%%%%
% CAPTIONS %
%%%%%%%%%%%%

%%%% Paragraph separation
%\setlength{\parskip}{.5em}

\numberwithin{equation}{section} % Number equations within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{figure}{section} % Number figures within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{table}{section} % Number tables within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)

%%%%%%%%%%%%
% DOCUMENT %
%%%%%%%%%%%%

\begin{document}

\title{\doctitle}
~\\
{\center
{\large 
{\large
\projectNameLong
}
\\[10mm]
\userName
\\[20mm]
{A dissertation \\
 submitted in partial fulfillment of the\\ requirements for the degree of}
\\[5mm]
{Doctor of Philosophy}
\\[20mm]
{\center
\textbf{Reading Committee}\\
Aleksandr Aravkin (co-chair),\\
J. Nathan Kutz (co-chair), \\
James Burke, \\
Jonathan Wakefield (GSR)
}
\\[20mm]
Program Authorized to Offer Degree:
\\[5mm]
\department
\\[5mm]
\institution
}
\\[10mm]
©Copyright 2023\\
\userName

}

\newpage
{
\center 
\institution
\\[10mm]
\textbf{Abstract}
\\[10mm]
\projectNameLong
\\[10mm]
\userName
\\[10mm]
Chairs of the Supervisory Committee:\\
Aleksandr Aravkin and J. Nathan Kutz\\
\department
\\[10mm]
}

My abstract goes here. I will probably write it at the very end.

\tableofcontents

\listoftodos

\chapter{Introduction}
\include{introduction}

\chapter{MSR3: Sparse Relaxed Regularized Regression for Mixed-Effect Models}
\label{ch:msr3}
\include{msr3}

\chapter{PINODE: Physics-Informed Neural Ordinary Differential Equations}
\label{ch:pinode}
\include{pinode}

\input{cs}

\chapter{Conclusion}

\bibliographystyle{apalike}
\bibliography{bibliography}

\clearpage

\appendix
\chapter{Appendix}
\section{Acknowledgement}
\input{appendix/acknowledgements}

\section{Derivatives of Marginalized Log-likelihood for Linear Mixed Models}
\label{appendix:derivatives_of_lmm}
\input{appendix/derivatives_of_lmms}

\section{Derivation of Selected Proximal Operators from Table \ref{table:proxes}}
\label{appendix:proxes}
\input{appendix/proxes}

\section{Existence of Minimizers (Theorems~\ref{thm:basic existence} and~\ref{thm:basic existence2})}
\label{adx:basic existence}
\input{appendix/existence_of_solutions_for_lme}

\section{Lipschitz-constant for Likelihood of a Linear Mixed-Effects Model}
\label{appendix:lipschitz_constant}
\input{appendix/lipschitz_constant}

\section{Detailed Results from Simulation from Table \ref{table:comparison_of_algorithms}}
\label{appendix:detailed_comparison}
\begin{table}[t]
    \centering
    \resizebox{\columnwidth}{!}{\input{tables/performance_table_long_current}}
    \caption{Comparison of performance of algorithms}
    \label{table:detailed_comparison_of_algorithms}
\end{table}

\section{Description of Real-World Datasets}
\subsection{GBD Bullying Data}
\input{appendix/bullying_features_description}

\subsection{COVID-19 Contact Rate Forecasting Data}
\input{tables/covid_groups_description}
\input{tables/covid_coefficients}

\section{Detailed Design of Experiments for PINODE}
\input{appendix/pinode_appendix}

\end{document}