### 2023-06-08 appendix/acknowledgements.tex
[FINAL SUBMITTED] The very final version of my thesis. Compiles clean.

I would like to thank my research advisors -- Dr. Aleksandr Aravkin, and Dr. Nathan Kutz, for their excellent leadership, continuous mentorship, and support in my journey as a Ph.D. student. I would also like to thank my collaborators -- Dr. James V. Burke, and Dr. Steven Brunton for their outstanding contributions to this work. I am grateful to Dr. Hassan Mansour and Dr. Saleh Nabi -- my colleagues and collaborators from Mitsubishi Electrics Research Labs who guided me to success throughout my internship in Spring 2023 and beyond. I wholeheartedly thank Biraj Pandey, Ryan Vogt, and Natalie Wellen -- my fellow Ph.D. students and dear friends from the Department of Applied Mathematics at UW: you turned what could have been a daunting and isolating journey into a wonderful experience full of happy memories. And even more so I would like to thank my family: my mom Svetlana Sholokhova, my grandmother Raisa Zakirova, and my wife Alanna Sholokhova. You have been surrounding me with unbelievable amounts of support and love, from close and afar, and I would have never succeeded in this journey without you.